\vspace{-2cm}A generative model for urban firm networks

### 2022-05-18 Response_to_reviewers.tex

%### Major issues

\begin{enumerate}
%1.1 -

	\item \textit{The work is at the intersection of three strands of literature: a) spatial interaction models, b) network analysis, and c) preferential attachment. Yet this does not appear in the literature review. The work could gain more depth from a better discussion of the literature it is grounded in.}

	$\rightarrow$ Literature review and paper positioning was reworked and extended to reflect this intersection of disciplines (although preferential attachment is not as central in our study).
	%\comment{(All) Rewrite lit review. Include applications of gravity models to trade and migration.}

@@

	$\rightarrow$ Additional statistical models extended to account for zeros were fitted and commented. This aspect was better explained.
	%(JR) Hurdle and zero inflated models\comment{(JR) Discuss choice/possibly test other stat models}

	\medskip

%1.3 -

	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue. Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectoral proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model. Can the authors mention (and may be discuss) this issue?}

@@

	\medskip

%1.4 -

	\item \textit{Application: Two potentially problematic aspects: 1) the MSE on logs and the log MSE appear to be quite large to perform an accurate prediction (for an MSE on logs $\sim 5$, the average multiplicative error can roughly be expected to be around $\exp(5^0.5) \sim 9$, which is almost an order of magnitude of difference; the same quick calculation can be applied to the log MSE). Can the authors discuss this point? (May be using a mean absolute error instead of mean squared error is a more tangible estimate of the error?)}

	$\rightarrow$ The purpose of both the statistical analysis and the generative model is not prediction. Statistical analysis is a preliminary exploration to determine potential processes to be included in the generative model, while the latest is an explanatory model, i.e. aimed at explaining the role of different factors in the growth of the urban network (such as the one of urban hierarchy) and test hypothesis in a virtual laboratory. It is also used to prototype policy applications, and in that sense is potentially a decision-making model. A discussion on model function and the precise positioning of this model was added.
	%\todo{(JR) this is not a prediction model - possibly give values of abs error - discuss more model function}

@@

	$\rightarrow$ Self-connections within urban areas are not considered in the current version of the model. We added in the discussion this possibility and issues related to such an extension.

	\medskip

% 2.2 - 
	\item \textit{Could you please better justify the choice of indicators (Internationalization and Metropolization).}

	% check lit - which lit stream
	$\rightarrow$ The choice of indicators has been better justified and explained based on citations of previous literature, already using these indicators to capture important globalisation processes in a city-level network.

### 2022-05-18 S1_Text.tex
Update on Overleaf.

Based on \emph {AMADEUS} dataset, the most vulnerable economic sectors in the United Kingdom have been highlighted. The later are the ones due to be mostly affected by eventual future lock-downs of the international economy (international ownership links).

When considering the UK sectors generating the highest turnover in 2018 mostly dependent on foreign capital they own (out-going links from the UK) these are extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods. In terms of in-going links to the UK, activities of head offices and holding companies, sale of cars and manufacture of motor vehicles are sectors mostly dependent on foreign ownership coming from abroad.

@@

\newpage
When expressed in total number of links, activities mostly dependent on out-going links from the UK are activities of holding companies, of head offices and fund management activities. In terms of in-going links these are business and consultancy activities and renting and operating of own and leased real estate.

>>>>

This section provides details on the selection of economic sectors in the United Kingdom that contain considerable international ownership links, and could hence become vulnerable to shocks related to those links being compromised, such as protectionist strategies, new trading agreements, reshuffling of the international trading organization, etc.

Looking specifically at sectors in the UK whose turnover is mostly dependent on their ownership of foreign capital in 2018, these correspond to extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods. Ownership relationships correspond to out-going links, see Table~\ref{tab:uk_outgoing} for details.

On the other hand, the sectors in the UK that mostly dependent on foreign ownership coming from abroad correspond to activities of head offices and holding companies, sale of cars and manufacture of motor vehicles, and these are represented by in-going links to the UK, see Table~\ref{tab:uk_ingoing}.

@@

\newpage
When expressed in total number of links (Table~\ref{tab:uk_outgoinglinks} and Table~\ref{tab:uk_ingoinglinks}), the activities that mostly depend on out-going links from the UK are activities of holding companies, of head offices and fund management activities. In terms of in-going links, these are business and consultancy activities and renting and operating of own and leased real estate.

### 2022-05-18 S2_Text.tex
Update on Overleaf.

In the real model setup, socio-economic distance between countries $c_{ij}$ is constructed using fixed effects coefficients from the statistical model. In Table~1 in main text, we estimated several statistical models, including some with fixed effects by pairs of countries. The socio-cultural distance between two countries is then taken as the opposite of the fixed effect coefficient for this pair, for the full statistical model. Summary statistics for the 29x29 matrix (28 EU countries and Jersey which is in the database a distinct country from UK but has an important total turnover since many British companies are located there for fiscal advantage reasons) are shown below in Table~\ref{tab:fixedeff}. We observe that distances are rather localized but with large outliers, and an important proportion are not defined (339 out of 841), corresponding to couple of countries having no exchange at all in the dataset.

@@

%pscl::hurdle(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
%    log(sim), data = dz, dist = "poisson", zero.dist = "poisson")
%
%Pearson residuals:
%       Min         1Q     Median         3Q        Max 
%-9.195e+03 -1.373e-01 -7.627e-02 -4.204e-02  1.727e+04 
%
%Count model coefficients (truncated poisson with log link):
%                     Estimate Std. Error z value Pr(>|z|)    
%(Intercept)        -9.853e+00  1.345e-04  -73252   <2e-16 ***
%log(distance)      -2.875e-01  7.603e-06  -37812   <2e-16 ***
%log(from_turnover)  7.379e-01  4.708e-06  156720   <2e-16 ***
%log(to_turnover)    6.379e-01  4.060e-06  157119   <2e-16 ***
%log(sim)            4.550e-01  2.137e-05   21292   <2e-16 ***
%Zero hurdle model coefficients (censored poisson with log link):
%                     Estimate Std. Error z value Pr(>|z|)    
%(Intercept)        -20.525823   0.155298 -132.17   <2e-16 ***
%log(distance)       -0.747788   0.010991  -68.04   <2e-16 ***
%log(from_turnover)   0.665678   0.005630  118.23   <2e-16 ***
%log(to_turnover)     0.479214   0.005707   83.96   <2e-16 ***
%log(sim)             0.278290   0.022439   12.40   <2e-16 ***
%---
%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
%
%Number of iterations in BFGS optimization: 46 
%Log-likelihood: -2.927e+10 on 10 Df

% 0.1661997

%mean((dz$intweight-hurdle2$fitted)^2)
%[1] 6.119694e+12

% > AIC(hurdle2)
%[1] 58545461912

%pscl::zeroinfl(formula = intweight ~ log(distance) + log(from_turnover) + log(to_turnover) + 
%    log(sim), data = dz, dist = "poisson")
%
%Pearson residuals:
%       Min         1Q     Median         3Q        Max 
%-7.693e+00 -1.335e-01 -6.957e-02 -3.591e-02  1.760e+04 
%
%Count model coefficients (poisson with log link):
%                     Estimate Std. Error z value Pr(>|z|)    
%(Intercept)        -9.853e+00  1.345e-04  -73251   <2e-16 ***
%log(distance)      -2.875e-01  7.603e-06  -37810   <2e-16 ***
%log(from_turnover)  7.379e-01  4.708e-06  156719   <2e-16 ***
%log(to_turnover)    6.379e-01  4.060e-06  157120   <2e-16 ***
%log(sim)            4.550e-01  2.137e-05   21294   <2e-16 ***
%
%Zero-inflation model coefficients (binomial with logit link):
%                    Estimate Std. Error z value Pr(>|z|)    
%(Intercept)        22.046685   0.180605  122.07   <2e-16 ***
%log(distance)       0.881724   0.013134   67.13   <2e-16 ***
%log(from_turnover) -0.737441   0.006776 -108.83   <2e-16 ***
%log(to_turnover)   -0.521642   0.006540  -79.76   <2e-16 ***
%log(sim)           -0.330340   0.025360  -13.03   <2e-16 ***
%---
%Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 
%
%Number of iterations in BFGS optimization: 1 
%Log-likelihood: -2.927e+10 on 10 Df

% 0.1607349

%mean((dz$intweight-zinflpoisson2$fitted)^2)
%[1] 6.159803e+12

% AIC(zinflpoisson2)
%[1] 58545461691

@@

The procedure for the synthetic sector setup is the following:
\begin{itemize}
    \item Sectors distributions follow log-normal densities with most mass in $\left[0;1\right]$
    \item Large cities are more innovative and more diverse. Assuming that sectors are ordered by innovativity (the larger the sector index the larger the innovativity), this assumption is translated by taking a mode and variance of 0.5 for the largest city and a mode and variance of $1/K$ for the smallest, and a linear interpolation between the two. For each city, we define
    \[
    \log(e_i) = \frac{(\log(E_i) - \log(E_{imin}))}{(\log(E_{imax}) - \log(E_{imin}))} * (1/2 - 1/K) + 1/K
    \]
    \item Log-normal parameters $(\mu_i,\sigma_i)$ for each city are then fixed by $\mu_i - \sigma_i^2 = \log(e_i)$ and $-3 \sigma_i^2 - 2 \log(\exp(\sigma_i^2) - 1) = log(e_i)$
    \item $\sigma_i^2$ is the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 \log(\exp(X) - 1) - \log(e_i)$ and $\mu_i = \log(e_i) + \sigma_i^2$.
\end{itemize}

%solve for mu,sigma of the log-normal as a function of gdp
%       we aim at having a log-normal mostly in [0,1] such that
%       (mode,variance) =  1 / K for the smallest log(E_i) and = 1/2 for the largest log(E_i)
%        writing
%        e_i = (log(E_i) - log(E_imin))/(log(E_imax) - log(E_imin)) * (1/2 - 1/K) + 1/K
%        this yields
%         (1)  mu - sigma^2 = log(e_i)
%         (2) -3 sigma^2 - 2 log(exp(sigma^2) - 1) = log(e_i)
%        => sigma^2 is the unique positive root of f(X)=0 with f(X) = -3X - 2 ln(exp(X) - 1) - ln(e_i)

@@

%%%%%%%%%%%%%
\begin{table}[h!]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
 & F & T & F & T & F & T & F & T & F & T & F & T \\
 \hline
Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.009 & $4\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
Metropolisation & 0.02 & 0.1 & 0.02 & 0.2 & 0.002 & 0.1 & 0.001 & 0.09 & 0.2 & 0.6 & 0.3 & 0.6 \\
Modularity & 0.3 & 0.4 & 0.6 & 0.6 & 0.004 & 0.02 & $3\cdot 10^{-4}$ & 0.01 & 0.005 & 0.03 & 0.002 & 0.03 \\
Avg. com. size & 0.008 & 0.09 & 0.01 & 0.1 & 0.002 & 0.07 & 0.003 & 0.04 & 0.3 & 0.6 & 0.4 & 0.6 \\
Degree entropy & 0.006 & 0.02 & 0.003 & 0.02 & 0.006 & 0.03 & 0.008 & 0.02 & 0.5 & 0.5 & 0.5 & 0.5 \\
Weight entropy & 0.04 & 0.1 & 0.03 & 0.1 & 0.008 & 0.08 & 0.01 & 0.07 & 0.4 & 0.5 & 0.4 & 0.5 \\\hline
\end{tabular}
\end{adjustwidth}
\end{table}
%%%%%%%%%%%%%

>>>>

In the real model setup, the socio-economic distance between countries $c_{ij}$ is constructed using the fixed effects coefficients from the statistical model. In Table~1 in the main text, we estimated several statistical models, including some with fixed effects by pairs of countries. The socio-cultural distance between two countries is then taken as the opposite of the fixed effect coefficient for this pair, for the full statistical model. The summary statistics for the 29x29 matrix (28 EU countries and Jersey, which is in the database corresponds to a distinct country from the UK, although it has an important total turnover since many British companies are located there for fiscal advantage reasons) are shown below in the Table~\ref{tab:fixedeff}. We observe that distances are rather localised but with large outliers, and an important proportion are not defined (339 out of 841), corresponding to a couple of countries having no exchange at all in the dataset.

@@

@@

The procedure for the synthetic sector setup is the following:
\begin{itemize}
    \item Sectors distributions follow log-normal densities with most mass in $\left[0;1\right]$
    \item Large cities are more innovative and more diverse. Assuming that sectors are ordered by innovativity (the larger the sector index the larger the innovativity), this assumption is translated by taking a mode and variance of $1/2$ for the sector distribution of the largest city and a mode and variance of $1/K$ for the sector distribution of the smallest city. Intermediate cities have their sector distribution mode and variance following a linear interpolation between the two. For each city, we therefore define this value for mode and variance as $e_i$ by taking
    \[
    e_i = \frac{(\log(E_i) - \log(E_{imin}))}{(\log(E_{imax}) - \log(E_{imin}))} \cdot (1/2 - 1/K) + 1/K
    \]
    where $E_i$ is the economic size of city $i$, $E_{imin}$ size of smallest city and $E_{imax}$ size of the largest city;
    \item Log-normal parameters $(\mu_i,\sigma_i)$ for each city are then determined by using the formula for the mode and variance of a log-normal law; the mode is directly given by $\mu_i - \sigma_i^2 = \log(e_i)$, while the variance verifies $\log(\exp(\sigma_i^2) - 1) + 2\mu + \sigma^2  = log(e_i)$; using the first equation to have an expression for $\mu$ and substituting into the second, we obtain
    \[
    -3 \sigma_i^2 - 2 \log(\exp(\sigma_i^2) - 1) = log(e_i)
    \]
    \item $\sigma_i^2$ is thus obtained as the unique positive root of $f(X)=0$ with $f(X) = -3X - 2 \log(\exp(X) - 1) - \log(e_i)$ (this function has a derivative always negative on $\mathbb{R}+$, tends to $\infty$ when $X \rightarrow 0+$ and to $-\infty$ when $X \rightarrow \infty$); we finally compute $\mu_i$ as $\mu_i = \log(e_i) + \sigma_i^2$.
\end{itemize}

@@

%%%%%%%%%%%%%
\begin{table}[h!]
\begin{adjustwidth}{-2.25in}{0in}
\caption{Saltelli sensitivity indices, for indicators in rows and parameters in columns. We give for each pair the first order index (F) and the total order index (T).\label{tab:saltelli}}
\hspace{-1cm}\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
 & \multicolumn{2}{|c|}{$d_0$} & \multicolumn{2}{|c|}{$c_0$} & \multicolumn{2}{|c|}{$\gamma_S$} & \multicolumn{2}{|c|}{$\gamma_W$} & \multicolumn{2}{|c|}{$\gamma_O$} & \multicolumn{2}{|c|}{$\gamma_D$} \\
 & F & T & F & T & F & T & F & T & F & T & F & T \\
 \hline
Internationalisation & 0.2 & 0.3 & 0.7 & 0.7 & 0.001 & 0.008 & $5\cdot 10^{-4}$ & 0.007 & 0.03 & 0.04 & 0.02 & 0.04 \\
Metropolisation & 0.02 & 0.04 & 0.02 & 0.04 & 0.02 & 0.3 & 0.002 & 0.01 & 0.2 & 0.7 & 0.2 & 0.7 \\
Modularity & 0.3 & 0.4 & 0.6 & 0.6 & $5\cdot 10^{-4}$ & 0.02 & 0.002 & 0.02 & 0.007 & 0.03 & 0.003 & 0.03 \\
Avg. com. size & 0.004 & 0.1 & 0.01 & 0.1 & 0.003 & 0.07 & 0.001 & 0.04 & 0.3 & 0.6 & 0.3 & 0.6 \\
Degree entropy & 0.003 & 0.03 & 0.002 & 0.03 & 0.008 & 0.04 & 0.003 & 0.03 & 0.5 & 0.6 & 0.5 & 0.6 \\
Weight entropy & 0.04 & 0.2 & 0.03 & 0.2 & 0.002 & 0.1 & 0.001 & 0.1 & 0.5 & 0.6 & 0.5 & 0.6 \\\hline
\end{tabular}
\end{adjustwidth}
\end{table}
%%%%%%%%%%%%%

### 2022-05-18 main.tex
Update on Overleaf.

\\
Juste Raimbault\textsuperscript{1,2,3 *},
Natalia Zdanowska\textsuperscript{3,4,5},
Elsa Arcaute\textsuperscript{4}
\\
\bigskip
\textbf{1} LASTIG, Univ Gustave Eiffel, IGN-ENSG, Saint-Mand{\'e}, France
\\
\textbf{2} UPS CNRS 3611 ISC-PIF, Paris, France
\\
\textbf{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
\\
\textbf{4} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
\\
\textbf{5} Luxembourg Institute of Socio-Economic Research, Esch-sur-Alzette, Luxembourg

@@

%The first main contribution of this paper is to focus on interurban interactions between firms. Firm networks remains a central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective. In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary} \cite{berry1964cities}. An extensive literature exists to explore firm-to-firm level networks in economic geography withing the input-output models framework in order to study inter-sectors and inter-region transactions \cite{Rodriguesetal2016} or firm level productivity to understand macroeconomic level outcomes \cite{Grassi2016} and economic resilience \cite{Ialongo2021}. Studies of these firm networks at city level remain still scarce. Urban firm networks have been investigated in complexity science to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.

%The second contribution of the present study is to question and capture geographical and economic processes as internationalisation and metropolisation of cities in terms of firm networks within a generative model, where weights of cities are given by shares of turnovers generated by ownership links between firms. Previous literature considers mostly weights as GDP of countries for example in gravity models of export-import firms networks \cite{Kepaptsogloual2010} \cite{MinondoRequena2013}. The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}. In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}. Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

@@

%The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study. We then perform a sensitivity analysis and exploration of the model parameters using synthetic data. The model is then calibrated on empirical European firm ownership data. We finally discuss implications of our results and possible extensions.

@@

To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we fit extended statistical models for count data focused on the role of zeros. Two different flavors of such models are: the hurdle model \cite{mullahy1986specification}, which corresponds to a conditional count model; and a zero-inflated Poisson model \cite{lambert1992zero}, which mixes a count model with an additional component to generate zeros. Full estimation results using \cite{zeileis2008regression}, are shown in S2 Text. As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links are difficult to predict. The best zero-inflated model has a $R^2 = 0.16$, and it is similar to the performance of the best hurdle moel $R^2 = 0.17$.

@@

A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators. We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}. This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y_j\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters. This captures the total influence of $X_i$ all other things being equal. The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters. Indices were estimated with a Sobol sequence sampling of 20,000 model runs \cite{saltelli2010variance}. We give in S2 Text the full estimated results for sensitivity indices. To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters. On the contrary, indicators of network structure are driven by origin and destination sizes. The influence of sectors and historical links has a secondary but not negligible role. Altogether, the analysis shows strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index. This means that the generative model captures some aspects of the complex behaviour of the system.

@@

The calibration of the model suggests its potential application to policy issues and can therefore have practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}). The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.

>>>>

\\
Juste Raimbault\textsuperscript{1,2,3,4 *},
Natalia Zdanowska\textsuperscript{2,4,5},
Elsa Arcaute\textsuperscript{2}
\\
\bigskip
\textbf{1} LASTIG, Univ Gustave Eiffel, IGN-ENSG, Saint-Mand{\'e}, France
\\
\textbf{2} Centre for Advanced Spatial Analysis, UCL, London, United Kingdom
\\
\textbf{3} UPS CNRS 3611 ISC-PIF, Paris, France
\\
\textbf{4} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s, Paris, France
\\
\textbf{5} Luxembourg Institute of Socio-Economic Research, Esch-sur-Alzette, Luxembourg

@@

@@

@@

To investigate the role of missing links in these results (these have been removed since previous models are fitted on logarithms), we fit extended statistical models for count data focused on the role of zeros. Two different flavors of such models are: the hurdle model \cite{mullahy1986specification}, which corresponds to a conditional count model; and a zero-inflated Poisson model \cite{lambert1992zero}, which mixes a count model with an additional component to generate zeros. Full estimation results using \cite{zeileis2008regression}, are shown in S2 Text. As the network is rather sparse, these more detailed statistical models do not perform better than more simple models, as missing links are difficult to predict. The best zero-inflated model has a $R^2 = 0.16$, and it is similar to the performance of the best hurdle model $R^2 = 0.17$.

@@

A global sensitivity analysis first unveils an aggregated influence of parameters on model indicators. We apply Saltelli's Global Sensitivity method introduced by \cite{saltelli2008global}. This produces values for each parameter and each output of a first order sensitivity index, defined as $Var\left[\mathbb{E}_{\mathbf{X}_{\sim i}}\left(Y_j | X_i\right)\right] / Var\left[Y_j\right]$, for parameter $X_i$ and output $Y_j$, $\mathbf{X}_{\sim i}$ being all other parameters. This captures the total influence of $X_i$ all other things being equal. The total order index is given by $\mathbb{E}_{\mathbf{X}_{\sim i}} \left[Var(Y_j | \mathbf{X}_{\sim i})\right] / Var\left[Y_j\right]$ and captures non-homogeneity in behaviour and interaction between parameters. Indices were estimated with a Sobol sequence sampling of 20,000 parameter points (corresponding to 160,000 model runs) \cite{saltelli2010variance}. We give in S2 Text the full estimated results for sensitivity indices. To summarize, we find that indicators related to modularity (internationalisation and optimal modularity) are mostly influenced by geographical parameters. On the contrary, indicators of network structure are driven by origin and destination sizes. The influence of sectors and historical links has a secondary but not negligible role. Altogether, the analysis shows strong interaction effects between parameters, since for several indicators the total order index is one order of magnitude larger than the first order index. This means that the generative model captures some aspects of the complex behaviour of the system.

@@

The model calibration results provide model parametrisations reasonably approaching the observed network. This validates its potential application to stylised policy issues with practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in detail are: the extraction of crude petroleum, the manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; the activities of head offices and holding companies, the sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalisation and thus can be expected to have a detrimental effect on the UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}). The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.
